A thought-experiment for a Planckian standard kilogram

### 2017-07-04 manuscript_gravitationalconstant.tex
definitions with numerical factors
writing intro

\title{Could a more precise measurement of the gravitational constant lead to Planckian SI units?}
% In a long title you can use \\ to force a line break at a certain location.

\author{Prasenjit Saha}
\email{psaha@physik.uzh.ch}
% If there were a second author at the same address, we would put another 
% \author{} statement here.  Don't combine multiple authors in a single
% \author statement.
\author{Michael Ebersold}
\email{michael.ebersold@uzh.ch} 
% Please provide a full mailing address here.
\affiliation{Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland}

@@

\section{Introduction} % Section titles are automatically converted to all-caps.
% Section numbering is automatic.

In the late 19th century, Max Planck was working on reports about the irreversible process of radiation. He worked out a new entropy law to explain the empirical Wien law of radiation. This required him to introduce two new physical constants, which nowadays are known as the Planck constant and the Boltzmann constant \cite{Tomilin1999}. As a consequence, he described in 1899 a unit system which is based on fundamental natural constants only and thus free of anthropocentric arbitrariness. He pointed out the advantage of such a system in the following words:\cite{Planck99}\\

@@

To come away from the definitions based on artefacts, specific physical states and idealized experiments and because it is technologically possible to measure physical constants to high accuracy, the International Committee for Weights and Measures (CIPM) has proposed new definitions of the SI base units, which likely will be adopted in fall 2018 at the next General Conference on Weights and Measures. The revision of the SI unit system in 2018 goes towards a Planckian based system, since the values of seven physical constants get fixed and thereby define the seven base units. 
To get a truly Planckian, hence ``most natural", basis for the SI, one would have to exchange the ground state hyperfine splitting frequency of the caesium-133 atom $\Delta \nu_{\mathrm{Cs}}$ with the gravitational constant $G$ and then define the second by  
\begin{equation}
\mathrm{s} = \sqrt{\frac{h \, G}{c^5}}.
\end{equation}
Because these three natural constants would have fixed numerical values, this definition is exact. Such a definition would lead to an SI system that is based on fundamental natural constants and therefore stable over time and realizable everywhere. Also completely new and more precise measurements could be designed without having to revise the definitions of the units \cite{Fischer16}. The big problem is, that the uncertainty in the measurements of $G$ is very large large. In Fig.~\ref{fig:Gmeasurements} the latest and most precise measurements of the gravitational constant are plotted together with the recommended CODATA value and its uncertainty. At first glance one notices, that the measured values are not even in good agreement with each other.\\

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{img/plotGmeas.eps}
	\caption{The data points represent six of the most recent measurement of the gravitational constant. The green line represents the recommended CODATA value and its uncertainty.}
	\label{fig:Gmeasurements}
\end{figure}

>>>>

\title{A thought-experiment for a Planckian standard kilogram}
% In a long title you can use \\ to force a line break at a certain location.

\author{Michael Ebersold}
\email{michael.ebersold@uzh.ch} 
% If there were a second author at the same address, we would put another 
% \author{} statement here.  Don't combine multiple authors in a single
% \author statement.
\author{Prasenjit Saha}
\email{psaha@physik.uzh.ch}
% Please provide a full mailing address here.
\affiliation{Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland}

@@

\section{Introduction}

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For an simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[citation].

The proposed redefinition of the kilogram is as an explicit value for
$h$ and $c$, but we can think of it as
\begin{equation}
\mathrm{kg} \equiv 1.3563925\ldots \times 10^{50} \, \mathrm{s}^{-1}
                   \times \frac{h}{c^2}
\end{equation}
with the exact value to be decided.  As we can see, the kilogram is
coupled to the definition of a second.  That is not a problem in
practice, as the second is standardized orders of magnitude better
than any other unit.  Nonetheless, it is interesting to consider a
logical further step of decoupling the kilogram from other
definitions.  This could be done by introducing the gravitational
constant, since $\sqrt{hc/G}$ has dimensions of mass.  Historically,
there have been several proposed unit systems that included $G$ in
some way,\cite{Tomilin1999} but the idea of including both gravity and
quantum in a unit system comes from Planck\cite{Planck99}.  In the
process of working towards the radiation law, Planck realized that
$c,G,h$ and the Boltzmann constant $k_B$ (in his notation
$G\rightarrow f$, $h\rightarrow b$, $h/k_B\rightarrow a$) naturally
define units for mass, length, time, and temperature.  (Planck
remarked that such units would be meaningful even in non-human and
extraterrestrial cultures.)

\begin{equation}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}

In the late 19th century, Max Planck was working on reports about the irreversible process of radiation. He worked out a new entropy law to explain the empirical Wien law of radiation. This required him to introduce two new physical constants, which nowadays are known as the Planck constant and the Boltzmann constant . As a consequence, he described in 1899 a unit system which is based on fundamental natural constants only and thus free of anthropocentric arbitrariness. He pointed out the advantage of such a system in the following words:\cite{Planck99}\\

@@

To come away from the definitions based on artefacts, specific physical states and idealized experiments and because it is technologically possible to measure physical constants to high accuracy, the International Committee for Weights and Measures (CIPM) has proposed new definitions of the SI base units, which likely will be adopted in fall 2018 at the next General Conference on Weights and Measures. The revision of the SI unit system in 2018 goes towards a Planckian based system, since the values of seven physical constants get fixed and thereby define the seven base units. 
To get a truly Planckian, hence ``most natural", basis for the SI, one would have to exchange the ground state hyperfine splitting frequency of the caesium-133 atom $\Delta \nu_{\mathrm{Cs}}$ with the gravitational constant $G$ and then define the second by  
\begin{equation}
\mathrm{s} = \sqrt{\frac{h \, G}{c^5}}.
\end{equation}

Because these four natural constants would have fixed numerical values, this definition is exact. Such a definition would lead to an SI system that is based on fundamental natural constants and therefore stable over time and realizable everywhere. Also completely new and more precise measurements could be designed without having to revise the definitions of the units \cite{Fischer16}. The big problem is, that the uncertainty in the measurements of $G$ is very large large. In Fig.~\ref{fig:Gmeasurements} the latest and most precise measurements of the gravitational constant are plotted together with the recommended CODATA value and its uncertainty. At first glance one notices, that the measured values are not even in good agreement with each other.\\

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{img/plotGmeas}
	\caption{The data points represent six of the most recent measurement of the gravitational constant. The green line represents the recommended CODATA value and its uncertainty.}
	\label{fig:Gmeasurements}
\end{figure}
